---
title: "Mediation Regression, 1980-2010"
output: html_notebook
---
```{r, echo = FALSE}
library(dplyr)
library(mediation)
library(fastDummies)
```
### Data Processing
```{r}
old <- read.csv("psid_old_stem.csv")
psid_old <- subset(old, select = c(sex, famwgt, age, sch, white, south, LEHS, black, hisp, othrace, west, northeast, northcentral, annhrs, realhrwage))
psid_old <- psid_old %>% mutate(annincome = realhrwage * annhrs) # calculate annual salary/wage
psid_old <- subset(psid_old, select = -c(realhrwage, annhrs, othrace))
psid_old$sex <- 1 - psid_old$sex

psid_old_male <- subset(psid_old, psid_old$sex == 1, select = -c(sex,northcentral))
psid_old_female <- subset(psid_old, psid_old$sex == 0, select = -c(sex,northcentral))
```

### Are X variables influencing wage?
Yes, sex(1 = Male), lower level of education(LEHS) reduces salary, while as age grows, the salary increases. These are highly significant, whereas Black people receiving less salary.  
```{r}
model.0.old.m <- lm(annincome~ . - sch, data = psid_old_male)
summary(model.0.old.m)

model.0.old.f <- lm(annincome~ . - sch, data = psid_old_female)
summary(model.0.old.f)
```

### Are X influencing school years?
Yes, mediation of school years make sense.  
```{r}

model.M.old.m <- lm(sch ~ . -annincome, data = psid_old_male)
summary(model.M.old.m)


model.M.old.f <- lm(sch ~ . -annincome, data = psid_old_female)
summary(model.M.old.f)
```

### Is school years affecting wages, without interfering X ?
Partial Mediation, not purely. We still have significant X variables, but they are smaller compared to the model above. As beta for school year(sch) is 1648 is also significant, we say this relationship exists.  
```{r}
model.Y.old.m <- lm(annincome ~ . , data = psid_old_male)
summary(model.Y.old.m)

model.Y.old.f <- lm(annincome ~ . , data = psid_old_female)
summary(model.Y.old.f)
```

### Is the mediation effect significant? - Bootstrap
ACME (Average Causal Mediation Effects)   
ADE (Direct Effect)  
Yes, the lower bond exceeds 0, meaning at 95% confidence level, we do believe education brings positive increase in wages, with sex variable significant. 
```{r}
psid_old_male <- subset(psid_old, psid_old$sex == 1, select = -c(sex,northcentral))
psid_old_female <- subset(psid_old, psid_old$sex == 0, select = -c(sex,northcentral))
bb <- 1000
nn_2 <-  min(nrow(psid_old_male),nrow(psid_old_female))
ACME <- matrix(0, bb, 1)				# new output matrix for R^2
CME.f.1 <- matrix(0, bb, 1)
CME.m.1 <- matrix(0, bb, 1)
# Do Data Bootstrap 1000 times to get 95% CI for R^2
for(ii in 1:bb) {
	
	data.m <- psid_old_male[sample(nn_2, nn_2, replace = T),]		# create (y*, x*) by resampling rows in original data matrix
  model.M.m <- lm(sch ~ . -annincome, data = data.m)							# save rsq from iteration ii
	model.Y.m <- lm(annincome ~ . , data = data.m)
	CME.m <- sum(model.M.m$coefficients) * model.Y.m$coefficients['sch']
	CME.m.1[ii] <- CME.m
	
	data.f <- psid_old_female[sample(nn_2, nn_2, replace = T),]		# create (y*, x*) by resampling rows in original data matrix
  model.M.f <- lm(sch ~ . -annincome, data = data.f)							# save rsq from iteration ii
	model.Y.f <- lm(annincome ~ . , data = data.f)
	CME.f <- sum(model.M.f$coefficients) * model.Y.f$coefficients['sch']
	ACME[ii] <- CME.m - CME.f
	CME.f.1[ii] <- CME.f
}

quantile(ACME, probs = c(0.025, 0.975), na.rm=TRUE) 
```
```{r}
quantile(CME.m.1, probs = c(0.025, 0.975), na.rm=TRUE) 
quantile(CME.f.1, probs = c(0.025, 0.975), na.rm=TRUE) 
```
## Plot
```{r}
CI.m.1 <- (60112.87-22800.74)/2
CI.f.1 <- (23036.135 +9685.777) /2
ACME.1 <- (59702.33-10024.49)/2
mean.m.1 <- mean(CME.m.1)
mean.f.1 <- mean(CME.f.1)
mean.acme.1<-mean(ACME)

plot.df <- data.frame(Mean.Education.Effect = c(mean.m.1, mean.f.1, mean.acme.1), CI = c(CI.m.1, CI.f.1, ACME.1),
                      Category = c("Male","Female","Difference"))
p<-ggplot(plot.df, aes(x=Mean.Education.Effect, y=Category, color = Category)) + 
  geom_point()+
  geom_errorbar(aes(xmin=Mean.Education.Effect-CI, xmax=Mean.Education.Effect+CI))+
  labs(title="Boostrap Confidence Interval of ACME, 1980 - 2010")+
theme_minimal()
  
p
```


```{r}
ACME.df <- as.data.frame(ACME)
p1 <- ggplot(ACME.df, aes(x=V1))+geom_histogram(color="darkblue", fill="lightblue", bins = 100) +
      geom_vline(xintercept=14419.03, colour = "darkblue",linetype="dotted")+
      geom_vline(xintercept=22215.36, colour = "darkblue",linetype="dotted")+
      #geom_text(aes(x=14419.03, label="Lower Bound", y=50), colour="dark grey")+
      #geom_text(aes(x=22215.36, label="Upper Bound", y=50), colour="dark grey")+
      #geom_rect(aes(xmin = -Inf , xmax = 14419.03 , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
      #geom_rect(aes(xmin = 22215.36, xmax = Inf , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
     labs(title="Boostrap Distribution of ACME", x ="Mediation Effect Level", y = "Count")+
     theme_minimal()+
     theme(plot.title = element_text(hjust = 0.5),panel.grid.major = element_blank(),panel.grid.minor = element_blank())
     #theme_minimal(plot.title = element_text(hjust = 0.5))
p1
p2 <- ggplot(ACME.df, aes(x=V1)) +
      geom_density(color="darkblue", fill="lightblue")+
      geom_vline(xintercept=14419.03, colour = "darkblue",linetype="dotted")+
      geom_vline(xintercept=22215.36, colour = "darkblue",linetype="dotted")+
      labs(title="Boostrap Distribution of ACME", x ="Mediation Effect Level", y = "Count")+
      theme_minimal()+
      theme(plot.title = element_text(hjust = 0.5),panel.grid.major = element_blank(),panel.grid.minor = element_blank())
  

p2

p3 <- ggplot(ACME.df, aes(x=V1)) +
      geom_density(color="darkblue")+
      geom_vline(xintercept=14419.03, colour = "darkblue",linetype="dotted")+
      geom_vline(xintercept=22215.36, colour = "darkblue",linetype="dotted")+
      labs(title="Boostrap Distribution of ACME", x ="Mediation Effect Level", y = "Count")+
      theme_minimal()+
      theme(plot.title = element_text(hjust = 0.5),panel.grid.major = element_blank(),panel.grid.minor = element_blank())
p3  

p4 <- ggplot(ACME.df, aes(x=V1, y = ..density..))+geom_histogram( fill="lightblue", bins = 50) +
      geom_density(color="darkblue")+
      geom_vline(xintercept=14419.03, colour = "darkblue",linetype="dotted")+
      geom_vline(xintercept=22215.36, colour = "darkblue",linetype="dotted")+
      #geom_text(aes(x=14419.03, label="Lower Bound=14419.03", y=50), colour="dark grey", angle = 90)+
      #geom_text(aes(x=22215.36, label="Upper Bound", y=50), colour="dark grey")+
      #geom_rect(aes(xmin = -Inf , xmax = 14419.03 , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
      #geom_rect(aes(xmin = 22215.36, xmax = Inf , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
     labs(title="Boostrap Distribution of ACME, 1980-2010", x ="Mediation Effect Level", y = "Density")+
     theme_minimal()+
     theme(plot.title = element_text(hjust = 0.5),panel.grid.major = element_blank(),panel.grid.minor = element_blank(), rect = element_rect(fill = "transparent")) +
      geom_text(aes(x=14419.03, label="2.5% quantile = 14419.03", y=0.00019,family="serif", fontface="italic"), colour="darkblue", angle=90, size=3.5)+
      geom_text(aes(x=22215.36, label="97.5% quantile = 22215.36", y=0.00019,family="serif", fontface="italic"), colour="darkblue", angle=90, size=3.5)
     

p4

    
```


