---
title: "R Notebook"
output: html_notebook
---

# 1. data cleaning

```{r}
getwd()
library(readxl)
data = read_excel("J303421.xlsx")
```

Variable Labels

ER66001 RELEASE NUMBER\
ER66002 2017 FAMILY INTERVIEW (ID) NUMBER\
ER66017 AGE OF REFERENCE PERSON\
ER66018 SEX OF REFERENCE PERSON\
ER66195 BC20 MAIN OCC FOR JOB 1: 2010 CODE (RP) ER66217 BC33 HOURLY REGULAR RATE\
ER66231 BC41 YRS PRES EMP (RP-E)\
ER67046 G13 WAGES/SALARY-REFERENCE PERSON\
ER70882 L40 RACE OF REFERENCE PERSON-MENTION 1\
ER70893 L44 WTR GRADUATED HS-RP\
ER70944 L71 #YR WRKED FULLTIME-RP\
ER71227 REFERENCE PERSON WORK WEEKS-2016\
ER71233 REF PERSON TOTAL HOURS OF WORK-2016\
ER71530 CURRENT REGION\
ER71538 COMPLETED ED-RP

ER72001 RELEASE NUMBER\
ER72002 2019 FAMILY INTERVIEW (ID) NUMBER\
ER72017 AGE OF REFERENCE PERSON\
ER72018 SEX OF REFERENCE PERSON\
ER72195 BC20 MAIN OCC FOR JOB 1: 2010 CODE (RP) ER72217 BC33 HOURLY REGULAR RATE\
ER72231 BC41 YRS PRES EMP (RP-E)\
ER73069 G13 WAGES/SALARY-REFERENCE PERSON\
ER76897 L40 RACE OF REFERENCE PERSON-MENTION 1\
ER76908 L44 WTR GRADUATED HS-RP\
ER76962 L71 #YR WRKED FULLTIME-RP\
ER77249 REFERENCE PERSON WORK WEEKS-2018\
ER77255 REF PERSON TOTAL HOURS OF WORK-2018\
ER77591 CURRENT REGION\
ER77599 COMPLETED ED-RP

```{r}
data_2017 <- na.omit(data[0:15])
data_2017$year <- "2017"
data_2019 <- na.omit(data[16:30])
data_2019$year <- "2019"

colnames(data_2017) <- c(
  "RELEASE_NUMBER",                         
  "INTERVIEW_ID",      
  "AGE",                
  "SEX",                
  "OCC_2010",
  "HOURLY.REGULAR.RATE",               
  "YRS.PRES.EMP",              
  "WAGES",      
  "RACE", 
  "WTR.GRADUATED",                
  "YR.WRKED.FULLTIME",            
  "WORK.WEEKS",       
  "TOTAL.HOURS",    
  "CURRENT.REGION",                         
  "COMPLETED.ED",
  "YEAR"
)
colnames(data_2019) <- colnames(data_2017)
data_new <- rbind(data_2017, data_2019)
dim(data_new)
```

## create dummies

```{r}
library('fastDummies')
df <- dummy_cols(data_new, select_columns = c('SEX','RACE', 'CURRENT.REGION',"YEAR"),
             remove_selected_columns = TRUE)
names(df)
```

## drop rows with DK or refuse to answer

```{r}
library(dplyr)
df = filter(df, RACE_9 == 0 & AGE != 999 
            & YR.WRKED.FULLTIME != 98 & YR.WRKED.FULLTIME != 99 
            & WAGES != 9999998 & WAGES != 9999999 
            & WTR.GRADUATED != 9 & COMPLETED.ED != 99)
```

## join stem code to create stem and no-stem tables

```{r}
stem_code = read.csv("2010-census-stem-related-and-non-stem-occupation-code-list.csv")
colnames(stem_code) = c("occupation_des","OCC_2010","soc_code")
library(dplyr)
data_new = df %>% left_join(stem_code,by="OCC_2010")
```

```{r}
library(dplyr)
data_not_stem = filter(data_new, is.na(occupation_des))
data_stem = df %>% inner_join(stem_code,by="OCC_2010")
# write.csv(data_stem,"/Users/yutian.lei/Desktop/MSBA hackthon/stem_data.csv", row.names = FALSE)
# install.packages('fastDummies')
```

# 2. linear with interaction

## 2.1 stem occupation

```{r}
df_1 <- subset(data_stem,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))
df_1$year_female <- df_1$RACE_2*df_1$YEAR_2019
df_1$WAGES <- log(df_1$WAGES+1)
model <- lm(WAGES~.,data = df_1)
summary(model)
```

## 2.2 not stem occupation

```{r}
df_2 <- subset(data_not_stem,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))
df_2$year_female <- df_2$RACE_2*df_2$YEAR_2019
df_2$WAGES <- log(df_2$WAGES+1)
model2 <- lm(WAGES~.,data = df_2)
summary(model2)
```

## 2.3 overall occupation

```{r}
data_new$isstem <- ifelse(is.na(data_new$occupation_des),0,1)
df_3 <- subset(data_new,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))

df_3$stem_female <- df_3$RACE_2*df_3$isstem
df_3$year_female <- df_3$RACE_2*df_3$YEAR_2019
df_3$WAGES <- log(df_3$WAGES+1)

model3 <- lm(WAGES~.,data = df_3)
summary(model3)
```

# 3. bootstrap difference of $\beta$

Do Data Boostrap 1000 times to get 95% beta difference

```{r}
set.seed(5)

bb <- 400						# number of resampling
coef.out <- matrix(0, bb, 2)			# matrix to save rsq from bootstrap
  
for(ii in 1:bb) {
  rr <- model$resid
  nn <- length(rr)
  data.star <- df_1[sample(nn, nn, replace = T),]
  out.star <- lm(WAGES~.,data = data.star)							
  coef.female <- summary(out.star)$coefficients[7,1]
  
  rr2 <- model2$resid
  nn2 <- length(rr2)
  data.star2 <- df_2[sample(nn2, nn2, replace = T),]
  out.star2 <- lm(WAGES~.,data = data.star2)							
  coef.female2 <- summary(out.star2)$coefficients[7,1]
  
	coef.out[ii,1] <- coef.female
	coef.out[ii,2] <- coef.female2
	
	colnames(coef.out) <- c("STEM", "Not_STEM")
	}

```

```{r}
coef.out <- as.data.frame(t(coef.out))
coef.out$diff = coef.out$STEM - coef.out$Not_STEM
```

```{r}
(diff <-  quantile(coef.out[,c("diff")], probs = c(0.1, 0.5, 0.9)))
```

**insights**

discrimination on wages of gender 对tech行业更严重（stem vs no.stem）
