---
title: "R Notebook"
output: html_notebook
---

# 1. data cleaning

```{r}
getwd()
library(readxl)
data = read_excel("J303421.xlsx")
```

Variable Labels

ER66001 RELEASE NUMBER\
ER66002 2017 FAMILY INTERVIEW (ID) NUMBER\
ER66017 AGE OF REFERENCE PERSON\
ER66018 SEX OF REFERENCE PERSON\
ER66195 BC20 MAIN OCC FOR JOB 1: 2010 CODE (RP) ER66217 BC33 HOURLY REGULAR RATE\
ER66231 BC41 YRS PRES EMP (RP-E)\
ER67046 G13 WAGES/SALARY-REFERENCE PERSON\
ER70882 L40 RACE OF REFERENCE PERSON-MENTION 1\
ER70893 L44 WTR GRADUATED HS-RP\
ER70944 L71 #YR WRKED FULLTIME-RP\
ER71227 REFERENCE PERSON WORK WEEKS-2016\
ER71233 REF PERSON TOTAL HOURS OF WORK-2016\
ER71530 CURRENT REGION\
ER71538 COMPLETED ED-RP

ER72001 RELEASE NUMBER\
ER72002 2019 FAMILY INTERVIEW (ID) NUMBER\
ER72017 AGE OF REFERENCE PERSON\
ER72018 SEX OF REFERENCE PERSON\
ER72195 BC20 MAIN OCC FOR JOB 1: 2010 CODE (RP) ER72217 BC33 HOURLY REGULAR RATE\
ER72231 BC41 YRS PRES EMP (RP-E)\
ER73069 G13 WAGES/SALARY-REFERENCE PERSON\
ER76897 L40 RACE OF REFERENCE PERSON-MENTION 1\
ER76908 L44 WTR GRADUATED HS-RP\
ER76962 L71 #YR WRKED FULLTIME-RP\
ER77249 REFERENCE PERSON WORK WEEKS-2018\
ER77255 REF PERSON TOTAL HOURS OF WORK-2018\
ER77591 CURRENT REGION\
ER77599 COMPLETED ED-RP

```{r}
data_2017 <- na.omit(data[0:15])
data_2017$year <- "2017"
data_2019 <- na.omit(data[16:30])
data_2019$year <- "2019"

colnames(data_2017) <- c(
  "RELEASE_NUMBER",                         
  "INTERVIEW_ID",      
  "AGE",                
  "SEX",                
  "OCC_2010",
  "HOURLY.REGULAR.RATE",               
  "YRS.PRES.EMP",              
  "WAGES",      
  "RACE", 
  "WTR.GRADUATED",                
  "YR.WRKED.FULLTIME",            
  "WORK.WEEKS",       
  "TOTAL.HOURS",    
  "CURRENT.REGION",                         
  "COMPLETED.ED",
  "YEAR"
)
colnames(data_2019) <- colnames(data_2017)
data_new <- rbind(data_2017, data_2019)
dim(data_new)
```

## create dummies

```{r}
library('fastDummies')
df <- dummy_cols(data_new, select_columns = c('SEX','RACE', 'CURRENT.REGION',"YEAR"),
             remove_selected_columns = TRUE)
names(df)
```

## drop rows with DK or refuse to answer

```{r}
library(dplyr)
df = filter(df, RACE_9 == 0 & AGE != 999 
            & YR.WRKED.FULLTIME != 98 & YR.WRKED.FULLTIME != 99 
            & WAGES != 9999998 & WAGES != 9999999 
            & WTR.GRADUATED != 9 & COMPLETED.ED != 99)
```

## join stem code to create stem and no-stem tables

```{r}
stem_code = read.csv("2010-census-stem-related-and-non-stem-occupation-code-list.csv")
colnames(stem_code) = c("occupation_des","OCC_2010","soc_code")
library(dplyr)
data_new = df %>% left_join(stem_code,by="OCC_2010")
```

```{r}
library(dplyr)
data_not_stem = filter(data_new, is.na(occupation_des))
data_stem = df %>% inner_join(stem_code,by="OCC_2010")
```

```{r}
# write.csv(data_stem,"/Users/yutian.lei/Desktop/MSBA hackthon/stem_data.csv", row.names = FALSE)
# write.csv(data_not_stem,"/Users/yutian.lei/Desktop/MSBA hackthon/data_not_stem.csv", row.names = FALSE)
# install.packages('fastDummies')
```

# 2. linear with interaction

## 2.1 stem occupation

```{r}
df_1 <- subset(data_stem,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))
df_1$year_female <- df_1$RACE_2*df_1$YEAR_2019
df_1$WAGES <- log(df_1$WAGES+1)
model <- lm(WAGES~.,data = df_1)
summary(model)
```

ln(Wages) = 7.160498 -0.324637*Female + X ln(Wages) = 2.308438 -0.060361*Female + X

## 2.2 not stem occupation

```{r}
df_2 <- subset(data_not_stem,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))
df_2$year_female <- df_2$RACE_2*df_2$YEAR_2019
df_2$WAGES <- log(df_2$WAGES+1)
model2 <- lm(WAGES~.,data = df_2)
summary(model2)
```

## 2.3 overall occupation

```{r}
data_new$isstem <- ifelse(is.na(data_new$occupation_des),0,1)
df_3 <- subset(data_new,select = -c(RELEASE_NUMBER,INTERVIEW_ID,OCC_2010,soc_code,
                              HOURLY.REGULAR.RATE,occupation_des,YR.WRKED.FULLTIME,
                              TOTAL.HOURS,
                              # omit dummy
                              SEX_1,RACE_1,RACE_9,CURRENT.REGION_1,YEAR_2017))

df_3$stem_female <- df_3$RACE_2*df_3$isstem
df_3$year_female <- df_3$RACE_2*df_3$YEAR_2019
df_3$WAGES <- log(df_3$WAGES+1)

model3 <- lm(WAGES~.,data = df_3)
summary(model3)
```

ln(Wages) = 2.328242 -0.068868\*Female + X

# 3. bootstrap difference of $\beta$

Do Data Boostrap 1000 times to get 95% beta difference

```{r}
set.seed(33)
bb <- 1000						# number of resampling
coef.out <- matrix(0, bb, 2)			# matrix to save rsq from bootstrap
  
for(ii in 1:bb) {
  rr <- model$resid
  nn <- length(rr)
  data.star <- df_1[sample(nn, nn, replace = T),]
  out.star <- lm(WAGES~.,data = data.star)							
  coef.female <- summary(out.star)$coefficients[7,1]
  
  rr2 <- model2$resid
  nn2 <- length(rr2)
  data.star2 <- df_2[sample(nn2, nn2, replace = T),]
  out.star2 <- lm(WAGES~.,data = data.star2)							
  coef.female2 <- summary(out.star2)$coefficients[7,1]
  
	coef.out[ii,1] <- coef.female
	coef.out[ii,2] <- coef.female2
	
	colnames(coef.out) <- c("Tech", "Overall")
	}

```

## tech and overall

```{r}
set.seed(33)
bb <- 1000						# number of resampling
coef.out <- matrix(0, bb, 2)			# matrix to save rsq from bootstrap
  
for(ii in 1:bb) {
  rr <- model$resid
  nn <- length(rr)
  data.star <- df_1[sample(nn, nn, replace = T),]
  out.star <- lm(WAGES~.,data = data.star)							
  coef.female <- summary(out.star)$coefficients[7,1]
  
  rr2 <- model3$resid
  nn2 <- length(rr2)
  data.star2 <- df_3[sample(nn2, nn2, replace = T),]
  out.star2 <- lm(WAGES~.,data = data.star2)							
  coef.female2 <- summary(out.star2)$coefficients[7,1]
  
	coef.out[ii,1] <- coef.female
	coef.out[ii,2] <- coef.female2
	
	colnames(coef.out) <- c("Tech", "Overall")
	}
```


```{r}
df = as.data.frame(coef.out)
df$diff = df$Tech - df$Overall
(ci_90 <-  quantile(df[,c("diff")], probs = c(0.05, 0.5, 0.95)))

(tech_ci_90 <-  quantile(df[,c("Tech")], probs = c(0.05, 0.5, 0.95)))
(overall_ci_90 <-  quantile(df[,c("Overall")], probs = c(0.05, 0.5, 0.95)))
```
```{r}
data <- data.frame(
  indus = c("Tech", "Overall","Diff"),
  mean = c(-0.32214849,-0.06772328,-0.2605556533),
  lower = c(-0.59022685,-0.12105137,-0.5279592189),
  upper = c(-0.12105137,-0.01409971,-0.0007971169)
)
 
# Most basic error bar
ggplot(data) +
    geom_bar(aes(x=indus, y=mean), stat="identity", fill="skyblue", alpha=0.7) +
    geom_errorbar( aes(x=indus, ymin=lower, ymax=upper), width=0.4, colour="orange", alpha=0.9, size=1.3) +
  xlab("Industry")+ylab("the Mean of Coeffcient of Is_Female")+
  theme_bw()+
  theme_classic()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  ggtitle("Boostrap Confidence Interval of Coeffcients")
```

```{r}
p4 <- ggplot(df, aes(x=diff, y = ..density..))+geom_density(color="darkblue")+
  geom_histogram(fill="lightblue", bins = 50) +
      geom_vline(xintercept=-0.0007971169, colour = "darkblue",linetype="dotted")+
      geom_vline(xintercept=-0.5279592189, colour = "darkblue",linetype="dotted")+
      #geom_text(aes(x=14419.03, label="Lower Bound=14419.03", y=50), colour="dark grey", angle = 90)+
      #geom_text(aes(x=22215.36, label="Upper Bound", y=50), colour="dark grey")+
      #geom_rect(aes(xmin = -Inf , xmax = 14419.03 , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
      #geom_rect(aes(xmin = 22215.36, xmax = Inf , ymin = -Inf, ymax = Inf),fill = "#FFFFCC", alpha = 0.015)+
     labs(title="Boostrap Distribution of Difference", x ="Difference of Female Coeffcients", y = "Density")+
     theme_minimal()+
     theme(plot.title = element_text(hjust = 0.5),panel.grid.major = element_blank(),panel.grid.minor = element_blank()) 
      # geom_text(aes(x=-0.5280634, label="5% quantile = -0.5280634", y=2,family="serif", fontface="italic"), colour="darkblue", angle=90, size=3.5)+
      # geom_text(aes(x=-0.007716132, label="95% quantile = -0.007716132", y=2,family="serif", fontface="italic"), colour="darkblue", angle=90, size=3.5)
     
p4
```

```{r}
library(reshape2)
mdata <- melt(df)
levels(mdata$variable)
names(mdata) = c("IsTech","Coeff")
mdata$IsTech <- as.factor(mdata$IsTech)

library(dplyr)
n = 1000
my_sum <- mdata %>%
  group_by("IsTech") %>%
  summarise( 
    mean=mean(Coeff),
    sd=sd(Coeff)
  ) %>%
  mutate(se=sd/sqrt(n))  %>%
  mutate(ic=se * qt((1-0.1)/2 + .5, n-1))
```

```{r}
my_sum
```


```{r}
ggplot(mdata) +
  geom_bar(aes(x=IsTech, y=Coeff), stat="identity", fill="skyblue", alpha=0.5) +
  geom_crossbar(aes(x=IsTech, y=Coeff, ymin=Coeff-sd, ymax=Coeff+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)
```

```{r}

```

```{r}
levels(mdata$variable)
```


```{r}
library(tidyr)

mdata<-gather(df)
names(mdata) = c("IsTech","Coeff")
```

```{r}
ggplot(data) +
  geom_bar(aes(x=name, y=value), stat="identity", fill="skyblue", alpha=0.5) +
  geom_crossbar(aes(x=name, y=value, ymin=value-sd, ymax=value+sd), width=0.4, colour="orange", alpha=0.9, size=1.3)
```


```{r}
library(plyr)
mu <- ddply(mdata, "IsTech", summarise, grp.mean=mean(Coeff))

library(ggplot2)
# Change density plot line colors by groups
p <- ggplot(mdata, aes(x=Coeff, color=IsTech)) +
  geom_density()+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=IsTech),
             linetype="dashed")
p
```



```{r}
factor(mdata$IsTech)
```


```{r}

```

```{r}
my_sum
```

```{r}
library(ggplot2)
ggplot(my_sum) +
  geom_bar(aes(x=IsTech, y=mean), stat="identity", fill="skyblue", alpha=0.7) +
  geom_errorbar(aes(x=IsTech, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, size=1.3) +
  ggtitle("using confidence interval")
```




```{r}
ACME.df <- as.data.frame(coef.out)
ACME.df
```

```{r}
ACME.df$diff = ACME.df$STEM - ACME.df$Not_STEM
```

```{r}
df <- as.data.frame(coef.out)

library(reshape2)
mdata <- melt(df)
names(mdata) = c("isstem","coeff")

```



```{r}
(ci_90 <-  quantile(ACME.df[,c("diff")], probs = c(0.05, 0.5, 0.95)))
ci_90[3]
```

```{r}

```

**insights**

discrimination on wages of gender 对tech行业更严重（stem vs no.stem）
